---
title: "Differential Analysis"
output: 
  html_document: 
    number_sections: yes
    theme: cosmo
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- import_biom(BIOMfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/otu_table_mc10_w_tax_rdp.biom',
                         treefilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.tre',
                         refseqfilename = '/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/otus/rep_set.fna',
                        refseqFunction = parse_taxonomy_default
                      )

mappingfile <- read.delim('/shares/CIBIO-Storage/CM/mir/projects/fbeghini_16s/20180731_run186ms_segata/grandi/mappingfile', row.names = 1, check.names = FALSE)
mappingfile$Description %<>% str_replace_all(.,'_','-')
mappingfile$condizione <- factor(mappingfile$condizione, labels = c('Pentatome + Bifido', 'Pentatome + PBS', 'Empty + Bifido', 'Empty + PBS'))
sample_names(gg_topi) <- mappingfile[sample_names(gg_topi), 'Description']
rownames(mappingfile) <- mappingfile$Description

sample_data(gg_topi) <- mappingfile
sample_data(gg_topi)$id.topo <- factor(sample_data(gg_topi)$id.topo)

sample_data(gg_topi)$pentatome <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,1])
sample_data(gg_topi)$bifido <- factor(stringr::str_split(sample_data(gg_topi)$condizione, ' \\+ ', simplify = TRUE)[,2])

sample_data(gg_topi)$cond_all <- factor(with(sample_data(gg_topi), paste(timepoint,real_vaccine, real_bifido, sep = '_')))
sample_data(gg_topi)$cond_vac <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_vaccine, sep = '_')))
sample_data(gg_topi)$cond_bif <- factor(with(sample_data(gg_topi), paste(merge_timepoint,real_bifido, sep = '_')))

sample_data(gg_topi)$vacc_bif <- strtoi(paste(ifelse(sample_data(gg_topi)$real_vaccine=='YES','1','0'),ifelse(sample_data(gg_topi)$real_bifido=='YES','1','0'), sep = ''), 2)

colnames(tax_table(gg_topi)) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")

tax_table_filtered <- tax_table(gg_topi) %>%
  as.data.frame %>% 
  mutate_all( .funs = function(x) stringr:::str_replace_all(x,"D_.__",""))

rownames(tax_table_filtered) <- rownames(tax_table(gg_topi))
tax_table(gg_topi) <-  as.matrix(tax_table_filtered)

unass_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(Domain == 'Unassigned') %>%
                select(OTU)

unknown_otus <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Genus)) %>%
                select(OTU)

unknown_family <- tax_table(gg_topi) %>%
                data.frame %>%
                tibble::rownames_to_column('OTU') %>%
                filter(is.na(Family)) %>%
                select(OTU)


tax_table(gg_topi)[unass_otus$OTU, ] <- rep.int('Unassigned', 7)
tax_table(gg_topi)[unknown_otus$OTU, 'Genus'] <- 'Unknown'
tax_table(gg_topi)[unknown_family$OTU, 'Family'] <- 'Unknown'

gg_topi.relab <- transform_sample_counts(gg_topi, function(x) x/sum(x))

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.family <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.family.relab <- transform_sample_counts(gg_topi.family, function(x) x/sum(x))
```


# Differential Analysis with DESeq2
```{r, cache=TRUE}
threshold <- 0.05

ggtopi_filt <- filter_taxa(gg_topi, function(x) mean(x) > 5e-1, TRUE)

pds <- phyloseq_to_deseq2(ggtopi_filt, design = ~timepoint)
dds <- DESeq2::DESeq(pds, parallel = TRUE, quiet = TRUE)

pds_genus <- phyloseq_to_deseq2(gg_topi.genus, design = ~timepoint)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE)

pds_family <- phyloseq_to_deseq2(gg_topi.family, design = ~timepoint)
dds_family <- DESeq2::DESeq(pds_family, parallel = TRUE, quiet = TRUE)
```

## T2 vs T3: Tumor vs no tumor injected
```{r}
res <- DESeq2::results(dds, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt <- res[!is.na(res$padj),]
res.filt <- res.filt[res.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_filt)[rownames(res.filt),])) %>%
  filter(baseMean >= 10) %>%
  ggplot(aes(log2FoldChange, interaction(Family,Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t23_samples_idx <- grep('T2|T3', sample_names(gg_topi.relab), perl = TRUE)
  
  otu_ab_t <- otu_table(gg_topi.relab)[rownames(res.filt), t23_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))
  

  
  otu_ab_t %>% 
    filter(OTU %in% (res.filt[res.filt$baseMean >= 10,] %>% rownames)) %>%
    bind_cols(Genus = as.matrix(tax_table(gg_topi.relab))[.$OTU,'Genus'],
              Family = as.matrix(tax_table(gg_topi.relab))[.$OTU,'Family']) %>%
    ggplot() +
      geom_boxplot(aes(interaction(Family, Genus, OTU), Abundance, fill=timepoint)) + 
      coord_flip()
```

## T2 vs T3: Tumor vs no tumor injected at Genus level
```{r}
res_genus <- DESeq2::results(dds_genus, contrast = c('timepoint', 'T2','T3'), alpha = 0.05)
res.filt_genus <- res_genus[!is.na(res_genus$padj),]
res.filt_genus <- res.filt_genus[res.filt_genus$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res.filt_genus %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.genus)[rownames(res.filt_genus),-7])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of Genus abundances of differenitally abundant Genera
```{r}
  t23_samples_idx_genus <- grep('T2|T3', sample_names(gg_topi.genus.relab), perl = TRUE)
  
  genus_ab_t <- otu_table(gg_topi.genus.relab)[rownames(res.filt_genus), t23_samples_idx_genus] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  genus_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## T2 vs T3: Tumor vs no tumor injected at Family level
```{r}
res_family <- DESeq2::results(dds_family, contrast = c('timepoint', 'T2','T3'), alpha = 0.05) 
res.filt_family <- res_family[!is.na(res_family$padj),]
res.filt_family <- res.filt_family[res.filt_family$padj<0.05,]
```

### Dotplot log2FC vs Family
```{r}
res.filt_family %>% as.data.frame %>% cbind(as.matrix(tax_table(gg_topi.family)[rownames(res.filt_family),-7])) %>%
  ggplot(aes(log2FoldChange, Family)) +
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t23_samples_idx_family <- grep('T2|T3', sample_names(gg_topi.family.relab), perl = TRUE)
  
  family_ab_t <- otu_table(gg_topi.family.relab)[rownames(res.filt_family), t23_samples_idx_family] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  family_ab_t %>% bind_cols(data.frame(tax_table(gg_topi.family.relab)[family_ab_t$OTU,c('Phylum','Family')])) %>%
  ggplot() +
    geom_boxplot(aes(Family, Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## T2 vs T1: Tumor vs no tumor injected
```{r}
res21 <- DESeq2::results(dds, contrast = c('timepoint', 'T2','T1'), alpha = 0.05)
res21.filt <- res21[!is.na(res21$padj),]
res21.filt <- res21.filt[res21.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res21.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_filt)[rownames(res21.filt),])) %>%
  filter(baseMean >= 10) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot abundances
```{r}
  t12_samples_idx <- grep('T2|T3', sample_names(ggtopi_filt), perl = TRUE)
  
  otu_ab_t12 <- otu_table(gg_topi.relab)[rownames(res21.filt), t12_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t12 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t12$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```


# T3 vs T7
```{r, cache=TRUE}
ggtopi_subs <-subset_samples(gg_topi, timepoint %in% c('T3','T7'))
ggtopi_subs.relab <- transform_sample_counts(ggtopi_subs, function(x) x/sum(x))
ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))
mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)

pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
```

## What is altered at T7 vs T3?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_subs), perl = TRUE)
  
  otu_ab_t37 <- otu_table(ggtopi_subs.relab)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_subs.relab)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 vs T3 at genus levels?
```{r}
res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),-7])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of Genus abundances of differenitally abundant Genera
```{r}
  genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after vaccination and Bifidobacterium gavage?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_vaccineYES.real_bifidoYES", alpha = 0.05) 
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)
  tax_table(gg_topi.relab)[rownames(res37.filt),]


  otu_ab_t37 <- otu_table(gg_topi.relab)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after vaccination only?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_vaccineYES", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs OTU
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs)[rownames(res37.filt),])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of OTUs abundances of differenitally abundant OTUs
```{r}
  t37_samples_idx <- grep('T7|T3', sample_names(ggtopi_filt), perl = TRUE)
  
  otu_ab_t37 <- otu_table(gg_topi.relab)[rownames(res37.filt), t37_samples_idx] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13))

  otu_ab_t37 %>% bind_cols(data.frame(tax_table(ggtopi_filt)[otu_ab_t37$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=timepoint)) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```

## What is altered at T7 after bifido gavage?
```{r}
res37 <- DESeq2::results(dds_37, name = "timepointT7.real_bifidoYES", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

NOTHING


# Removal of all Bacteroidales 24/7
```{r}
gg_topi_no247 <- subset_taxa(gg_topi, !Family %in% c("Bacteroidales S24-7 group"))
```

# T3 vs T7 without 24/7
```{r, cache=TRUE}
ggtopi_subs <-subset_samples(gg_topi_no247, timepoint %in% c('T3','T7'))
ggtopi_subs.genus <- tax_glom(ggtopi_subs,taxrank = "Genus")
ggtopi_subs.genus.relab <- transform_sample_counts(ggtopi_subs.genus, function(x) x/sum(x))
mmtx <- model.matrix(object = ~timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, data = data.frame(ggtopi_subs@sam_data))
idx <- which(apply(mmtx, 2, function(x) all(x==0)))
mmtx <- mmtx[,-idx]

pds_37 <- phyloseq_to_deseq2(ggtopi_subs, design = ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_37 <- DESeq2::DESeq(pds_37, parallel = TRUE, quiet = TRUE, full = mmtx)

pds_genus <- phyloseq_to_deseq2(ggtopi_subs.genus, design =  ~ timepoint + timepoint:real_vaccine + timepoint:real_bifido + timepoint:real_vaccine:real_bifido, ignoreRank = TRUE)
dds_genus <- DESeq2::DESeq(pds_genus, parallel = TRUE, quiet = TRUE, full = mmtx)
```

## What is altered at T7 vs T3 at genus levels?
```{r}
res37 <- DESeq2::results(dds_genus, name = "timepointT7", alpha = 0.05)
res37.filt <- res37[!is.na(res37$padj),]
res37.filt <- res37.filt[res37.filt$padj<0.05,]
```

### Dotplot log2FC vs Genus
```{r}
res37.filt %>% as.data.frame %>% 
  cbind(as.matrix(tax_table(ggtopi_subs.genus)[rownames(res37.filt),-7])) %>%
  ggplot(aes(log2FoldChange, interaction(Family, Genus))) + 
  geom_vline(xintercept = 0.0, color = "gray", size = 0.5) +
  geom_point(size=3, aes(color = baseMean)) + 
  facet_grid(Phylum ~ ., scales = "free", space = "free_y",switch = "y",as.table = T)+
  theme_bw() +
  scale_y_discrete(position = "right") +
  theme(axis.text = element_text(hjust = 0, vjust=0.5, size = 11),
        strip.text = element_text(size = 9),
        axis.title = element_text(size = 9),
        axis.title.y = element_blank(),
        strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 180)
  )
```

### Boxplot of Genus abundances of differenitally abundant Genera
```{r}
  genus_ab_t <- otu_table(ggtopi_subs.genus.relab)[rownames(res37.filt), ] %>%
              data.frame %>%
              tibble::rownames_to_column("OTU") %>%
              tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
              mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
              mutate(condition= if_else(timepoint=='T3','',stringr::str_sub(.$SampleID,4, 6)))
    
  
  genus_ab_t %>% bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(interaction(Family, Genus), Abundance, fill=interaction(timepoint, condition))) + 
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    coord_flip() + 
    theme_bw() +
    theme(strip.background = element_rect(colour = "transparent", fill = "transparent"),
        strip.text.y = element_text(angle = 0))
```