---
title: "GG_topi"
author: "Francesco Beghini"
date: "26/09/2018"
output: 
  html_document: 
    df_print: kable
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, paged.print=FALSE, echo=FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
library(phyloseq)
library(magrittr)
library(tidyverse)
library(reshape2)
library(ggsignif)
library(DESeq2)

source('R/loadQiimeData.R')
source('R/calculatePairwiseDistances.R')
```

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- loadQiimeData()

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

gg_topi.phylum <- tax_glom(gg_topi,taxrank = "Phylum")
gg_topi.phylum.relab <- transform_sample_counts(gg_topi.phylum, function(x) x/sum(x))
```
<!-- # Taxonomy profile at genus level -->
<!-- ```{r taxonomy, cache=FALSE} -->
<!-- # nychanesmicrobiome::plot_abundance(gg_topi,taxrank = 'Genus') -->
<!-- ``` -->

VVP: Vuote + PBS

VPP: Pentatoma + PBS

VVB: Vuote + *Bifidobacterium*

VPB: Pentatoma + *Bifidobacterium*

```{r Calculate alpha diversity}
alpha_div <- estimate_richness(gg_topi)
alpha_div %<>% bind_cols(data.frame(sample_data(gg_topi)))
```

# Alpha diversity: timepoint x Index x Condition
```{r Plot Alpha diversity boxplots}
melt(alpha_div, id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'se.chao1', 'Simpson', 'Shannon')) %>%
ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  facet_grid(variable~condizione,scales = 'free_y')
```

## Alpha diversity: tumor vs no tumor
```{r}
filt_alpha <- alpha_div %>% 
  filter(timepoint %in% c('T1', 'T2', 'T3')) %>%
  select('Description','condizione','timepoint','Observed', 'se.chao1', 'Simpson', 'Shannon') %>%
  rename('se.chao1' = 'Chao1') %>%
  mutate(t=as.numeric(stringr::str_sub(timepoint,2,2))) %>% droplevels()

ttest_res <- data.frame()
for ( ts in list(c('T1','T2'), c('T2','T3')) ){
  for( measure in c('Observed', 'Chao1', 'Simpson', 'Shannon') ){
    ttest_res %<>% bind_rows( data.frame(Measure = measure,
                                         Timepoint = paste(ts, collapse = ' vs '),
                                         pvalue = with(filt_alpha %>% filter(timepoint %in% ts), t.test(as.formula(paste(measure, 't', sep = '~'))))$p.value))
  }
}

filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value)) +
  xlab('Timepoint') +
  ylab('Value') +
  facet_grid(variable~.,scales = 'free_y') +
  theme_minimal() +
  theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.y = element_text(angle = 0))

knitr::kable(ttest_res)
```

# Beta diversity
```{r Beta diversity, cache=TRUE}
distwu <- phyloseq::distance(gg_topi, "wunifrac")
distbc <- phyloseq::distance(gg_topi, "bray")
distjac <- phyloseq::distance(gg_topi, "jaccard")

ordwu <-  ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordbc <-  ordinate(gg_topi, method = "MDS", distance = distbc, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)
ordjac <-  ordinate(gg_topi, method = "MDS", distance = distjac, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

distanceswu <- calculate_pairwise_distances(distwu)
distancesbc <- calculate_pairwise_distances(distbc)
distancesjac <- calculate_pairwise_distances(distjac)
```

## Ordination: Mean all samples per timepoint
```{r}
x <- ordinate(gg_topi, method = "MDS", distance = distwu, weighted = TRUE) %>% plot_ordination(gg_topi, ., justDF = TRUE)

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)

ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 0.8) +
  geom_point(data=centroids, size = 4) + 
  theme_minimal()
```


## Ordination: T2 vs T3, tumor vs no tumor
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T2','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T2','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),R^2==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = -.05, ymax = -.05)
```

### Merge T1 and T2 and use them as baseline
## Ordination: baseline vs T3
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T3'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T4
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T4'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T5
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T5'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T6
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T6'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: baseline vs T7
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('baseline','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=merge_timepoint, shape=real_vaccine), size=3) +
  theme_bw()
```

## Ordination: T3 vs T7: Treated with Bifidum?
```{r}
ggplot(subset(ordwu, merge_timepoint %in% c('T3','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=cond_all), size=3) +
  scale_colour_discrete(name="Timepoint_Vaccine_Bifido") +
  theme_bw() 
```
 
```{r}
ggplot(subset(ordwu, timepoint %in% c('T3','T7'))) +
  geom_point(aes(Axis.1, Axis.2, color=timepoint, shape=bifido), size=3) +
  facet_wrap(pentatome~bifido)+
  theme_bw() 
```


## Ordination: T3 vs T4
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T4','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T4','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ timepoint + real_vaccine + timepoint:real_vaccine, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T4','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint, shape=real_vaccine), size=3, alpha=.7) +
  scale_color_brewer(name = "Timepoint", palette="Dark2") + 
  scale_shape_manual(name = 'Pentatome?', values=c(15,19))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('Timepoint PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = .2, xmax = .2, ymin = .1, ymax = .1) +
  annotation_custom(grob = grid::textGrob(paste('Pentatome PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[2], '\nR2=', signif(permanova$aov.tab$R2[2], 2)), 
gp=grid::gpar(col="black", fontsize=10)),  xmin = -.1, xmax = -.1, ymin = -.08, ymax = -.08)
```

## Ordination: T3 vs T5
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T5','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T5','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T5','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .12, ymax = .12)
```

## Ordination: T3 vs T6
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T6','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T6','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T6','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Ordination: T3 vs T7
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(fill=timepoint, shape=real_vaccine, color=real_bifido), size=3, alpha=.7, stroke=0.8) +
  scale_fill_brewer(name = "Timepoint", palette="Dark2") +
  scale_shape_manual(name = 'Pentatome?', values=c(21,22))+
  scale_color_manual(name = 'Bifido?', values = c('blue','red'))+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```

## Beta diversity between timepoints, T1 is the baseline
```{r Beta diversity between timepoints t1}
distances_tmp <- distanceswu
distances_tmp[distances_tmp$timepoint1 == 1 | distances_tmp$timepoint1 == 2, 'timepoint1'] <- 2
distances_tmp[distances_tmp$timepoint2 == 1 | distances_tmp$timepoint2 == 2, 'timepoint2'] <- 2

t1 <- distances_tmp %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==2) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(t1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
    pval.t = t.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"])$p.value
    pval.w = wilcox.test(t1[t1$timepoint2==x[1] & t1$condizione1==cond,"value"], t1[t1$timepoint2==x[2] & t1$condizione1==cond ,"value"], paired = F)$p.value
    if(pval.t < 0.05 | pval.w < 0.05)
      t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
    }
  }

```

```{r}
knitr::kable(t.res)
```

```{r}

t1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  stat_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)
  
```

## Beta diversity between timepoints, T(-1) is the baseline
```{r Beta diversity between timepoints tn-1}
tn1 <- distanceswu %>% 
  filter(topo1==topo2) %>%
  filter(timepoint1==timepoint2-1) %>%
  filter(timepoint1!=timepoint2) %>%
  arrange(topo1)

t.res <- data.frame()
for(cond in unique(tn1$condizione1)){
  for(x in combn(c(3:7), 2, simplify = FALSE)){
      pval.t = t.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"])$p.value
      pval.w = wilcox.test(tn1[tn1$timepoint2==x[1] & tn1$condizione1==cond,"value"], tn1[tn1$timepoint2==x[2] & tn1$condizione1==cond ,"value"], paired = F)$p.value
      if(pval.t < 0.05 | pval.w < 0.05)
        t.res %<>% bind_rows(c(condizione1=cond, timepoint=as.numeric(x), p.value = str_c('t-test:',signif(pval.t, digits = 2),'U-test:', signif(pval.w, digits = 2), sep = ' ')))
  }
}
```

```{r}
knitr::kable(t.res)
```

```{r}
tn1 %>%
  ggplot(aes(timepoint2, value, group = timepoint2)) +
  geom_boxplot() +
  scale_x_continuous(breaks = c(2:7)) +
  # geom_signif(data=t.res,
              # aes(xmin= as.numeric(timepoint1), xmax = as.numeric(timepoint2), annotations = p.value, y_position = 0.33),
              # manual = TRUE) +
  facet_grid(condizione1 ~ .)
```

# PERMANOVA
```{r, cache=TRUE}
metadata <- as(sample_data(gg_topi), "data.frame")
permanova.res <- data.frame()
distwu <- phyloseq::distance(gg_topi, "wunifrac")

all_permanova <- as.data.frame(vegan::adonis(formula = distwu ~ merge_timepoint*bifido*pentatome, data = metadata)$aov.tab)
```

```{r}
knitr::kable(all_permanova)
```

```{r, cache=TRUE}
for(s in combn(levels(metadata$merge_timepoint),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$merge_timepoint %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```


```{r, cache=TRUE}
for(s in combn(levels(metadata$pentatome),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$pentatome %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, pentatome %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ pentatome, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```


```{r, cache=TRUE}
for(s in combn(levels(metadata$bifido),2, simplify = FALSE)){
  metadata_2cat <- metadata[metadata$bifido %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, bifido %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ bifido, data = metadata_2cat)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
knitr::kable(permanova.res)
```

## (T5+T6+T7) vs baseline
```{r}
metadata_tub <- metadata[metadata$merge_timepoint %in% c('baseline','T5','T6','T7'),]
metadata_tub$merge_timepoint <- as.character(metadata_tub$merge_timepoint)
metadata_tub[which(metadata_tub$merge_timepoint!='baseline'), 'merge_timepoint'] <- 'Timm'
distwu_tub <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('baseline','T5','T6','T7')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_tub ~ merge_timepoint, data = metadata_tub)
```

```{r}
knitr::kable(permanova$aov.tab)
```

## Effect of bifido in T4 vs T5
```{r}
metadata_t45 <- metadata[metadata$merge_timepoint %in% c('T4','T5'),]
distwu_t45 <- phyloseq::distance(subset_samples(gg_topi, merge_timepoint %in% c('T4','T5')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_t45 ~ merge_timepoint*bifido, data = metadata_t45)
```

```{r}
knitr::kable(permanova$aov.tab)
```

## Beta diversity T3 vs T4
```{r}
t1 <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==4) %>%
  dplyr::arrange(topo1)

t1 %>%
  ggplot(aes(pentatome1, value)) +
  stat_boxplot() +
  stat_signif()
```

## Beta diversity T3 vs T7
```{r}
t37 <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2==7) %>%
  dplyr::arrange(topo1)

t37 %>%
  ggplot(aes(condizione1, value)) +
  stat_boxplot()
```


## Correlation beta-diversity T3-TX, tumor mass
```{r corr tumor wunifrac}
distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
wunif_ds <- distanceswu %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_wunif <- data.frame()
for(x in unique(wunif_ds$timepoint2)){
  for(y in levels(wunif_ds$condizione)){
    tmp_res <- with(wunif_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_wunif %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_wunif)
```

```{r corr tumor braycurtis}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
brayc_ds <- distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_braycf <- data.frame()
for(x in unique(brayc_ds$timepoint2)){
  for(y in levels(brayc_ds$condizione)){
    tmp_res <- with(brayc_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_braycf %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_braycf)
```

```{r corr tumor jaccard}
distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint2~condizione, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
jac_ds <- distancesjac %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description'))

ds_corr_jac <- data.frame()
for(x in unique(jac_ds$timepoint2)){
  for(y in levels(jac_ds$condizione)){
    tmp_res <- with(jac_ds %>% filter(timepoint2==x & condizione == y), cor.test(value, tumor_mass))
    ds_corr_jac %<>% bind_rows(data.frame(Cond = paste(y, ' at T', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}

knitr::kable(ds_corr_jac)
```

```{r, include=FALSE, eval=FALSE}
otus <- data.frame(otu_table(gg_topi.genus.relab), check.names = FALSE) %>% 
  tibble::rownames_to_column('OTU') %>% 
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  tibble::add_column(., timepoint = stringr::str_sub(.$SampleID, 12, 13)) %>%
  tibble::add_column(., mouseID = stringr::str_sub(.$SampleID, 8, 10))

wb<-createWorkbook(type="xlsx")
for(mID in sort(unique(otus$mouseID))){
  tmp <- otus %>%
    filter(mouseID==mID) %>%
    reshape2::dcast(OTU ~ timepoint, value.var = 'Abundance') %>%
    left_join(data.frame(tax_table(gg_topi.genus.relab)[,-7], check.names = FALSE) %>% tibble::rownames_to_column('OTU'))
  sheet <- createSheet(wb, sheetName = mID)
  addDataFrame(tmp, sheet)
}
saveWorkbook(wb, "otutable_topi.xlsx")
```

# Distribution of distances between conditions per timepoint
```{r}
dct <- distanceswu %>%
  filter(timepoint1 == timepoint2) %>%
  select(timepoint1, condizione1, condizione2, value) %>%
  dplyr::rename(timepoint = timepoint1) %>% 
  group_by(timepoint, condizione1, condizione2) %>%
  summarise(mediandistance=median(value)) 
```

```{r include=TRUE,  results = "asis"}
for(t in unique(dct$timepoint)){
  print(dct %>%
    filter(timepoint == t) %>% 
    dplyr::rename(condizione = condizione2) %>%
    spread(condizione1, mediandistance) %>%
    knitr::kable())

}
```



# Distribution of distances between timepoints per conditions
```{r}
dtc <- distanceswu %>%
  filter(condizione1 == condizione2) %>%
  dplyr::rename(condizione = condizione1) %>% 
  group_by(condizione, timepoint1, timepoint2) %>%
  summarise(mediandistance=median(value)) 
```

```{r include=TRUE,  results = "asis"}
for(c in unique(dtc$condizione)){
  print(dtc %>%
    filter(condizione == c) %>% 
    dplyr::rename(timepoint = timepoint1) %>%
    spread(timepoint2, mediandistance) %>%
    knitr::kable())
    cat("\n")
}
```


## PERMANOVA all vs all
```{r, cache=TRUE}
permanova.res <- data.frame()
for (s in combn(levels(sample_data(gg_topi)$cond_all),2, simplify = FALSE)) {
  metadata_2cat <- data.frame(sample_data(gg_topi))[sample_data(gg_topi)$cond_all %in% s,]
  distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, cond_all %in% s), "wunifrac")
  permanova <- vegan::adonis(formula = distwu_2cat ~ cond_all, data = metadata_2cat, parallel = 2)
  permanova.res %<>% bind_rows(data.frame(contrast = stringr::str_c(s, collapse = " vs "), r2=permanova$aov.tab[1,5],pvalue=permanova$aov.tab[1,6])) %>% arrange(pvalue)
}
```

```{r}
knitr::kable(permanova.res)
```


# Abundance of Bacteroidales 24/4 through timepoints
```{r}
idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")

otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  ggplot() +
    geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
    facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T)+
    # coord_flip() + 
    theme_bw() +
    theme( strip.background = element_rect(colour = "transparent", fill = "transparent"),
           strip.text.y = element_text(angle = 0))
```

# Correlation of tumor size at T7 and abundance of Bacteroidales 24/ through timepoints
```{r}
mass_247 <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id-topo'))

mass_247 %>%
  ggplot(aes(Abundance,tumor_mass)) +
  geom_point() + 
  facet_grid(timepoint.x~condition, scales = "free", space = "free_y")+
  geom_smooth(method = 'lm')
```

```{r}
ds_corr_247 <- data.frame()
for(x in unique(mass_247$timepoint.x)){
  for(y in unique(mass_247$condition)){
    tmp_res <- with(mass_247 %>% filter(timepoint.x==x & condition==y), cor.test(Abundance, tumor_mass))
    ds_corr_247 %<>% bind_rows(data.frame(Cond = paste(y, ' ', x, sep = ''), corr = tmp_res$estimate, pval = tmp_res$p.value))
  }
}
print(knitr::kable(ds_corr_247))
```

# Correlation of tumor size at T7 and difference of abundance of Bacteroidales 24/7 through timepoints
```{r}
mass_247 %>%
  filter(timepoint.x %in% c('T3','T7')) %>%
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3)) %>%
  ggplot(aes(T3_7,tumor_mass)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

```{r}
with(mass_247 %>%
  filter(timepoint.x %in% c('T3','T7')) %>%
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3)),
  cor.test(T3_7,tumor_mass))

```

## Ordination: T3 vs T7 with tumor size
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(shape=timepoint, color=tumor_mass), size=3, alpha=.7, stroke=0.8) +
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(paste('PERMANOVA\np-value=',permanova$aov.tab$`Pr(>F)`[1], '\nR2=', signif(permanova$aov.tab$R2[1], 2)), gp=grid::gpar(col="black", fontsize=10)),  xmin = 0.18, xmax = .18, ymin = .1, ymax = .1)
```
